Reducing RES Droughts through the integration of wind and PV

### 2025-01-15 Article2/RES_droughts_Morin_RE.tex
Update Files Names

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

@@

%\usepackage{natbib}

@@

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

@@

%% use optional labels to link authors explicitly to addresses:
%% \author[label1, label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

@@

\cortext[cor1]{Correponding author}

@@

%%Research highlights
\begin{highlights}
\item The C3S-Energy dataset provides renewable energy capacity factor time series

\item RES droughts are analysed using 45 years of hourly wind and PV generation data

@@

The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \citep{eurostat2023share}. While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power. This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2021, bloomfield2016, vanderWiel2019drought}. Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.

@@

\begin{figure}[ht!]
	\centering
	\includegraphics{droughts_methodology.pdf}
	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021. The black dashed line indicates the CF threshold. The grey bar shows the period identified as a wind drought under our definition}
	\label{fig:find_res_droughts}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{verification_power_curve.pdf}
	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and Observed (shaded)}
	\label{fig:power_curve}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{verification_wind_contour.png}
	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N models}
	\label{fig:wind_verification_contour}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{verification_wind_number_events.pdf}
	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
	\label{fig:bar_number_events_verification_wind}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics{verification_pv_contour.png}
	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
	\label{fig:solar_verification_contour}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{verification_pv_number_events.pdf}
	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline). The PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
	\label{fig:bar_number_events_verification_pv}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{droughts_number_events.pdf}
	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events. The boxes display the first and third quartiles and the median is marked by a black line. The whiskers indicate the 5th and 95th percentiles}
	\label{fig:boxplot_number_events}	
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{droughts_return_periods.pdf}
	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square). The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it. The horizontal dashed line marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
	\label{fig:return_periods}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics{droughts_seasonality.pdf}
	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid). The x-axis represents the month of the year, and the y-axis indicates the percentage of hours. Lines correspond to the median values and the area between the first and third quartiles is shaded.}
	\label{fig:res_droughts_seasonality}
\end{figure}

@@

%% For citations use: 
%%       \citet{<label>} ==> Lamport (1994)
%%       \citep{<label>} ==> (Lamport, 1994)

@@

%% else use the following coding to input the bibitems directly in the
%% TeX file.

@@

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.

>>>>

%% The lineno packages adds line numbers.
\usepackage{lineno}

@@

@@

@@

@@

\cortext[cor1]{Corresponding author}

@@

%%Research highlights
\begin{highlights}
\item RES droughts are analysed using 45 years of hourly wind and PV generation data

@@

The EU aims to generate at least 69\% of its electricity from renewable energy sources (RES) by 2030, up from 41\% in 2022 \citep{eurostat2023share}. While this transition is essential for reducing greenhouse gas emissions, it also highlights the challenge of managing the variability of weather-dependent energy sources such as wind and photovoltaic (PV) power. This challenge is compounded by the increasing electrification of energy sectors, which places greater demand on the power system and makes it more sensitive to meteorological conditions~\citep{bloomfield2016, bloomfield2021, vanderWiel2019drought}. Periods of low renewable generation, known as \textit{Dunkelflaute} or RES droughts, pose significant risks to system adequacy and energy security, emphasizing the need for a resilient energy system to meet both growing electricity demand and decarbonization targets.

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{droughts_methodology.pdf}
	\caption{Wind time series of CF (green) and its 24-hour moving average (pink) from the 7th to the 15th of July 2021. The black dashed line indicates the CF threshold. The grey bar shows the period identified as a wind drought under our definition}
	\label{fig:find_res_droughts}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification_power_curve.pdf}
	\caption{a) Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (orange) and the Vestas V136/3450 used by C3S-E (teal) b) Histograms of wind CF for Ireland from Atlite (orange), C3S-E (teal) and Observed (shaded)}
	\label{fig:power_curve}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification_wind_contour.png}
	\caption{Wind CF density plot of the observed CF (vertical axes) and modelled (horizontal axes) CF data for the a) Atlite, b) C3S-E G and c) C3S-E N models}
	\label{fig:wind_verification_contour}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification_wind_number_events.pdf}
	\caption{Average annual number of wind drought events for Atlite (red), C3S-E G (blue), C3S-E N (purple), and the observed data (black outline). The wind droughts are identified from 2014 to 2023}
	\label{fig:bar_number_events_verification_wind}
\end{figure}

@@

\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{verification_pv_contour.png}
	\caption{PV CF density plot of the observed (vertical axes) and modelled (horizontal axes) CF series for the a) Atlite, b) C3S-E G and c) C3S-E N models}	
	\label{fig:solar_verification_contour}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification_pv_number_events.pdf}
	\caption{Number of PV drought events for Atlite (red), C3S-E G (blue), and C3S-E N (purple) and the observed data (black outline). The PV droughts are identified for 2023, considering the actual capacity of the system at any given time}
	\label{fig:bar_number_events_verification_pv}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_number_events.pdf}
	\caption{Annual number of RES droughts (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events. The boxes display the first and third quartiles and the median is marked by a black line. The whiskers indicate the 5th and 95th percentiles}
	\label{fig:boxplot_number_events}	
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_return_periods.pdf}
	\caption{Return periods of the duration of RES droughts  (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red triangle), C3S-E G (blue circle), and C3S-E N (purple square). The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it. The horizontal dashed line marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset}
	\label{fig:return_periods}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_seasonality.pdf}
	\caption{Percentage of hours in a month which are part of a RES drought (from 1979 to 2023) for a)~Wind, b)~PV, and the combination for the c)~91W-9PV and d)~57W-43PV scenario, for Atlite (red dotted), C3S-E G (blue dashed), and C3S-E N (purple solid). The x-axis represents the month of the year, and the y-axis indicates the percentage of hours. Lines correspond to the median values and the area between the first and third quartiles is shaded.}
	\label{fig:res_droughts_seasonality}
\end{figure}

@@

@@

@@

\endinput
